6CCS3PRJ \\\vspace{0.5cm} A Pragmatic Infrastructure Modelling and Verification Framework

### 2014-04-06 paper.tex
which -> that!

I discovered a real disparity between the knowledge application developers had of the specific requirements of their application (hardware specifications, dependencies on other services, likely points of failure) and the lack of awareness of this within the operations team. This phenomenon has been noted in other areas of the industry~\cite{DevOps} and has recently given rise to the term `DevOps' - an attempt to improve communication between the teams or build tools which help to do so.

@@

The objective is to design and implement a tool which helps with the modelling of complex infrastructure within real-world systems. The application will provide application developers or operations staff the ability to model systems as a graph of nodes, which may have dependencies and fine grained tests used to verify their current state.

@@

This model can serve as `living documentation' which is used by Ops teams and Developers to help manage and transfer knowledge, documenting their implicit knowledge. It should help to predict the impact of changes. It should improve collaboration between the teams and provide deeper insight into issues with the performance or reliability of the system.

Languages and tools for the modelling of software architecture and infrastructure do exist, however they lack the simplicity and pragmatism in their implementation to see widespread adoption in industry~\cite{ModellingAdoption}. I propose a simple, pragmatic tool which may come with some modelling limitations, but will be quick and easy to set up and begin reaping the benefits.

@@

The problem of increasing software complexity is a well publicised and researched phenomenon. Software which is complex takes longer to develop, and is harder to maintain. As it affects the value proposition to the user so directly, it is often the dominant factor in determining the feasibility of implementing a given system.

@@

The real value of these DSL's is that they are directly executable.  They can be executed to drive infrastructure changes and ensure convergence with the model. Figure~\ref{fig:PuppetDSLSnippet} provides an example snippet of a Puppet recipe, declaratively specifying details of the SSH (secure shell) service which will be made available on a specific host.

@@

I have extensively researched existing monitoring tools ranging from very simple ping checks to complete enterprise solutions and evaluated their effectiveness. In the interest of brevity, I have included a small sample of popular monitoring tools I have evaluated which give an indication of the cross-section of the market:

@@

Nagios is in active use by over 250,000 organisations worldwide, including Amazon, AT\&T and JP Morgan Chase~\cite{NagiosUsage}. It supports a number of common monitoring architectures, including both agent based and agent-less (based on the Simple Network Management Protocol) which means it can run efficiently up to a very large number of hosts~\cite{NagiosArchitectures}.

@@

This collection of tools all work together in providing near real-time metrics tracking and graphing, they are used heavily at vastly successful startups such as 37signals, GitHub and Shopify. Metrics can be collected from the infrastructure and application layers. 
CollectD is a system utility written in C which periodically makes system calls to extract performance fundamentals such as CPU, memory and network usage.

@@

This collection of tools provides an extremely flexible and performant source of aggregate data which can be used to diagnose reliability and performance problems effectively. Its disadvantages however include a lack of any alerting mechanism, a number of independent components which are both non-trivial to deploy and difficult to monitor themselves, and a difficulty in actually interpreting aggregate data to draw conclusions without deep insight into both the application and the infrastructure it’s relying on.

@@

There are a number of issues that prevent this process from being accurate in every scenario, but in practice it is a flexible solution which can detect most services accurately enough to be a useful time saving feature.

@@

\begin{itemize}
  \item Service - An application that provides functionality to an end-user or another service over a network\. e.g. a website or API.
  \item Component - Part of a service which performs a specific function, e.g.\ an application server or a load balancer.
  \item Host - A physical, virtual or otherwise isolated container responsible for running some or all of the components of a service.
  \item Metric - A measure of a characteristic at a given point in time, e.g.\ CPU load or average request latency. Typically quantitative in nature.
  \item Check - Validation that an arbitrary property holds at a given point in time, e.g.\ A specific service is listening on port 80. Typically qualitative in nature.
  \item Event - A discrete action performed upon the system, e.g.\ a code deployment or service re-configuration. Events can be correlated with metrics.
  \item Incident - A specific type of event which is automatically triggered by the failure of one or more checks, incidents correlate recent events with failed checks and can be related to previous incidents of a similar nature to aid in root-cause analysis.
  \item Context - Additional data which aids root-cause analysis, e.g.\ log files from a malfunctioning service.
\end{itemize}

@@

The Web UI will subscribe to a WebSocket connection to listen for changes in the system, this will ensure that any changes in the system are immediately visible to the end user, a feature which many other monitoring systems do not provide.

@@

  \item{method\_missing} - As a dynamically typed language, function calls in Ruby are dynamically dispatched depending on the type of an object. The mechanism for dynamic dispatch also includes a fallback feature used when there are no matching methods with the given name. This means that DSL's can be expressed which pattern match on method names or provide transparent proxies to other objects. Both of these features can be used extensively to create concise, readable DSL's.
\end{itemize}

@@

The most important feature of the DSL is the ability to encode checks as Ruby blocks. A block contains one or more assertions that throw exceptions with descriptive error messages if they fail. Below is a simple example of this. The health check will be executed every 30 seconds by the monitoring agent. The monitoring agent will catch any exceptions and return a descriptive error message which may be printed to the console or serialised and sent to the event stream.

@@

\begin{description}
  \item[Check] - Run configured health checks on demand. This command is typically used on developer workstations as they iteratively build a service specification using the DSL\. It provides fast feedback similar to running a suite of unit tests and can be used to quickly validate assumptions, before pushing the specification to the management server.
  \item[Discover] - Creates a service specification from detected system services. This feature can be used primarily to reverse-engineer existing infrastructure setups into a service specification. More details about the discovery mechanism are provided in the next subsection.
  \item[Graph] - Render a graph of the current service - This can be primarily used on the developer workstation, it will use the GraphViz library to render a graph of the system and its dependencies image or PDF format. This can be used on its own as a form of documentation, but is typically intended to help visualise the service specification and its dependencies to ensure its correctness.
  \item[Push] - Pushes the specified service to a management server - This command will parse the DSL and serialise each object to JSON, it is then sent as a PUT request to the management API\. Changes to the service specification take effect immediately and update any connected users in near real-time.
  \item[Setup] - Register this host with the management server - This is used on the servers running the system under test, it will associate a host with a defined service, listing it in the Web UI and accepting events and check data from it.
  \item[Start] - Starts a monitoring daemon process which runs in the background and executes checks according to their defined intervals. It will report all data back to the management server's event stream endpoint. Any errors occurring during this process will be logged to a file for later analysis.
  \item[Stop] - Stop monitoring the specified service.
  \item[Restart] - Restart monitoring of the specified service.
  \item[Help] - Describes the function of all available commands.
\end{description}

@@

The streaming event service is based on the open source project
Riemann\footnote{See http://riemann.io/concepts for an introduction to
Riemann.}. Riemann is a simple and scalable event processing framework written in Clojure\footnote{See http://clojure.org for a full description of Clojure}. Riemann simply provides a powerful set of primitive operations to filter, throttle and detect changes in streams of data. Riemann provides a configuration file, itself written in Clojure which is used to specify the desired behaviour.

@@

The streaming event service provides one endpoint to send all types of event data, this includes results from status checks, application performance metrics and log file entries. These types of event can all be processed and routed to a number of locations. This agnosticism makes it easy to adapt the monitoring architecture in future, hosts can be configured to send data to a single endpoint which in turn can route the events to their appropriate location.

@@

    Where this is particularly difficult due to a large number of side effects or dependencies, I will use integration testing to provide some high-level guarantees and ensure that the system satisfies its requirements. A small amount of manual testing will also be required.
  \item [Vagrant and Amazon AWS]\hfill \\
    Where integration tests are required, I will use Vagrant\footnote{See
    https://docs.vagrantup.com/v2/why-vagrant/index.html - Benefits of using Vagrant}, a Ruby scripting tool which controls
    VirtualBox\footnote{See https://www.virtualbox.org/ for an overview of VirtualBox} and automates the provisioning and setup of Virtual Machines according to a specification.

>>>>

I discovered a real disparity between the knowledge application developers had of the specific requirements of their application (hardware specifications, dependencies on other services, likely points of failure) and the lack of awareness of this within the operations team. This phenomenon has been noted in other areas of the industry~\cite{DevOps} and has recently given rise to the term `DevOps' - an attempt to improve communication between the teams or build tools that help to do so.

@@

The objective is to design and implement a tool that helps with the modelling of complex infrastructure within real-world systems. The application will provide application developers or operations staff the ability to model systems as a graph of nodes, which may have dependencies and fine grained tests used to verify their current state.

@@

This model can serve as `living documentation' that is used by Ops teams and Developers to help manage and transfer knowledge, documenting their implicit knowledge. It should help to predict the impact of changes. It should improve collaboration between the teams and provide deeper insight into issues with the performance or reliability of the system.

Languages and tools for the modelling of software architecture and infrastructure do exist, however they lack the simplicity and pragmatism in their implementation to see widespread adoption in industry~\cite{ModellingAdoption}. I propose a simple, pragmatic tool that may come with some modelling limitations, but will be quick and easy to set up and begin reaping the benefits.

@@

The problem of increasing software complexity is a well publicised and researched phenomenon. Software that is complex takes longer to develop, and is harder to maintain. As it affects the value proposition to the user so directly, it is often the dominant factor in determining the feasibility of implementing a given system.

@@

The real value of these DSL's is that they are directly executable.  They can be executed to drive infrastructure changes and ensure convergence with the model. Figure~\ref{fig:PuppetDSLSnippet} provides an example snippet of a Puppet recipe, declaratively specifying details of the SSH (secure shell) service that will be made available on a specific host.

@@

I have extensively researched existing monitoring tools ranging from very simple ping checks to complete enterprise solutions and evaluated their effectiveness. In the interest of brevity, I have included a small sample of popular monitoring tools I have evaluated that give an indication of the cross-section of the market:

@@

Nagios is in active use by over 250,000 organisations worldwide, including Amazon, AT\&T and JP Morgan Chase~\cite{NagiosUsage}. It supports a number of common monitoring architectures, including both agent based and agent-less (based on the Simple Network Management Protocol) that means it can run efficiently up to a very large number of hosts~\cite{NagiosArchitectures}.

@@

This collection of tools all work together in providing near real-time metrics tracking and graphing, they are used heavily at vastly successful startups such as 37signals, GitHub and Shopify. Metrics can be collected from the infrastructure and application layers. 
CollectD is a system utility written in C that periodically makes system calls to extract performance fundamentals such as CPU, memory and network usage.

@@

This collection of tools provides an extremely flexible and performant source of aggregate data that can be used to diagnose reliability and performance problems effectively. Its disadvantages however include a lack of any alerting mechanism, a number of independent components that are both non-trivial to deploy and difficult to monitor themselves, and a difficulty in actually interpreting aggregate data to draw conclusions without deep insight into both the application and the infrastructure it’s relying on.

@@

There are a number of issues that prevent this process from being accurate in every scenario, but in practice it is a flexible solution that can detect most services accurately enough to be a useful time saving feature.

@@

\begin{itemize}
  \item Service - An application that provides functionality to an end-user or another service over a network\. e.g. a website or API.
  \item Component - Part of a service that performs a specific function, e.g.\ an application server or a load balancer.
  \item Host - A physical, virtual or otherwise isolated container responsible for running some or all of the components of a service.
  \item Metric - A measure of a characteristic at a given point in time, e.g.\ CPU load or average request latency. Typically quantitative in nature.
  \item Check - Validation that an arbitrary property holds at a given point in time, e.g.\ A specific service is listening on port 80. Typically qualitative in nature.
  \item Event - A discrete action performed upon the system, e.g.\ a code deployment or service re-configuration. Events can be correlated with metrics.
  \item Incident - A specific type of event that is automatically triggered by the failure of one or more checks, incidents correlate recent events with failed checks and can be related to previous incidents of a similar nature to aid in root-cause analysis.
  \item Context - Additional data that aids in root-cause analysis, e.g.\ log files from a malfunctioning service.
\end{itemize}

@@

The Web UI will subscribe to a WebSocket connection to listen for changes in the system, this will ensure that any changes in the system are immediately visible to the end user, a feature that many other monitoring systems do not provide.

@@

  \item{method\_missing} - As a dynamically typed language, function calls in Ruby are dynamically dispatched depending on the type of an object. The mechanism for dynamic dispatch also includes a fallback feature used when there are no matching methods with the given name. This means that DSL's can be expressed that pattern match on method names or provide transparent proxies to other objects. Both of these features can be used extensively to create concise, readable DSL's.
\end{itemize}

@@

The most important feature of the DSL is the ability to encode checks as Ruby blocks. A block contains one or more assertions that throw exceptions with descriptive error messages if they fail. Below is a simple example of this. The health check will be executed every 30 seconds by the monitoring agent. The monitoring agent will catch any exceptions and return a descriptive error message, which may be printed to the console or serialised and sent to the event stream.

@@

\begin{description}
  \item[Check] - Run configured health checks on demand. This command is typically used on developer workstations as they iteratively build a service specification using the DSL\. It provides fast feedback similar to running a suite of unit tests and can be used to quickly validate assumptions, before pushing the specification to the management server.
  \item[Discover] - Creates a service specification from detected system services. This feature can be used primarily to reverse-engineer existing infrastructure setups into a service specification. More details about the discovery mechanism are provided in the next subsection.
  \item[Graph] - Render a graph of the current service - This can be primarily used on the developer workstation, it will use the GraphViz library to render a graph of the system and its dependencies image or PDF format. This can be used on its own as a form of documentation, but is typically intended to help visualise the service specification and its dependencies to ensure its correctness.
  \item[Push] - Pushes the specified service to a management server - This command will parse the DSL and serialise each object to JSON, it is then sent as a PUT request to the management API\. Changes to the service specification take effect immediately and update any connected users in near real-time.
  \item[Setup] - Register this host with the management server - This is used on the servers running the system under test, it will associate a host with a defined service, listing it in the Web UI and accepting events and check data from it.
  \item[Start] - Starts a monitoring daemon process, which runs in the background and executes checks according to their defined intervals. It will report all data back to the management server's event stream endpoint. Any errors occurring during this process will be logged to a file for later analysis.
  \item[Stop] - Stop monitoring the specified service.
  \item[Restart] - Restart monitoring of the specified service.
  \item[Help] - Describes the function of all available commands.
\end{description}

@@

The streaming event service is based on the open source project
Riemann\footnote{See http://riemann.io/concepts for an introduction to
Riemann.}. Riemann is a simple and scalable event processing framework written in Clojure\footnote{See http://clojure.org for a full description of Clojure}. Riemann simply provides a powerful set of primitive operations to filter, throttle and detect changes in streams of data. Riemann provides a configuration file, itself written in Clojure, which is used to specify the desired behaviour.

@@

The streaming event service provides one endpoint to send all types of event data, this includes results from status checks, application performance metrics and log file entries. These types of event can all be processed and routed to a number of locations. This agnosticism makes it easy to adapt the monitoring architecture in future, hosts can be configured to send data to a single endpoint, which in turn can route the events to their appropriate location.

@@

    Where this is particularly difficult due to a large number of side effects or dependencies, I will use integration testing to provide some high-level guarantees and ensure that the system satisfies its requirements. A small amount of manual testing will also be required.
  \item [Vagrant and Amazon AWS]\hfill \\
    Where integration tests are required, I will use Vagrant\footnote{See
    https://docs.vagrantup.com/v2/why-vagrant/index.html - Benefits of using Vagrant}, a Ruby scripting tool that controls
    VirtualBox\footnote{See https://www.virtualbox.org/ for an overview of VirtualBox} and automates the provisioning and setup of Virtual Machines according to a specification.
